---
title: "DATA_607_WEEK3"
author: "Keshaw Sahay"
date: "February 16, 2020"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    number_section:  true
    theme: cerulean
    highlight:  tango
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Problem 
Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"


```{r warning=FALSE, message=FALSE}
library(stringr)
library(tidyverse)
library(kableExtra)
library(knitr)
```

```{r}
major_df <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv")
kable(head(major_df))
```

```{r}
summary(major_df)
```

```{r}
reg_data_stats = str_detect(levels(major_df$Major), regex("DATA|STATISTICS", ignore_case=TRUE))
reg_data_stats
```
```{r}
levels(major_df$Major)[reg_data_stats]
```


#Problem 
Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"


```{r}
input_text <- '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"'


char_Vector <- c(unlist(str_extract_all(input_text, "\\b[A-Za-z]+\\b")))

vec_str <- str_c('"', char_Vector, '"', collapse = ", " )

final_text <- str_c('c(', vec_str, ')', collapse = " " )

#Final Output text
writeLines(final_text)
```

#Problem 

Describe, in words, what these expressions will match:

## (.)\\1\\1

```{r}
word_ex_char <- c("(.)\1\1", "(.)(.)\\2\\1", "(..)\1", "(.).\\1.\\1", "(.)(.)(.).*\\3\\2\\1")
word_exprs <- list("aaa", "abc", "abba", "afada", "ab\1", "a\1\1", "abccba")

str_view(word_exprs, '(.)\1\1')
```

This expression matches characters which are followed by a “\\1\\1”

## (.)(.)\\2\\1

```{r}
str_view(word_exprs, '(.)(.)\\2\\1')
```

It matches strings which contain pairs of characters that are followed by a reverse of their order.

## (..)\\1
```{r}
str_view(word_exprs, '(..)\1')
```

It matches strings a couple of characters that are followed by “\\1”. 

## (.).\\1.\\1

```{r}
str_view(word_exprs, '(.).\\1.\\1')
```
This expression matches strings a character that repeats in the 2 and 4 places over from its first occurrence.

## (.)(.)(.).*\\3\\2\\1

```{r}
str_view(word_exprs, '(.)(.)(.).*\\3\\2\\1')
```
This expression matches any sequence of strings of 3 characters, excluding line breaks, where the ending character is a reverse order of those 3 characters

# Problem 
Construct regular expressions to match words that

## Start and end with the same character

```{r}
word_1 <- list("blurb", "9Thousand9", "Light", "101DATA101", "MAGMA", "BANANA")
str_view(word_1, "^(.)(.*)\\1$") 
```

## Contain a repeated pair of letters (e.g. “church” contains “ch” repeated twice

```{r}
str_view(word_1, '([A-Za-z][A-Za-z]).*\\1')
```


## Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

```{r}
str_view(word_1, '([A-Za-z]).*\\1.*\\1')
```



