---
title: "DATA607 - Week 9 Assignment"
author: "Keshaw Sahay"
date: "March 29, 2018"
Dataset: "Working with Web APIs"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    number_section:  true
    theme: cerulean
    highlight:  tango
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. The New York Times web site provides a rich set of APIs, as described here: http://developer.nytimes.com/docs
You all need to start by signing up for an API key.
Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it to an R dataframe. 

```{r echo=FALSE,warning=FALSE, cache=FALSE, message=FALSE}
library(jsonlite)
library(rjson)
library(tidyverse)
library(RCurl)
library(DT)
apikey <- "GtPZ6gvZvQKtJ1Vkyq72iLAAVDEIbHDf"

```


# NYTimes web and developer API Key
Per NYTimes documentaion, we need to obtain the API key by registering as a developer at NYTimes developer webstite. 
<br> 
Getting Started: 
The Times Developer Network is our API clearinghouse and community. Here's how to get started:

* Request an API key
* Read the API documentation, FAQ and Terms of Use
* Use the API Tool associated with each API to experiment without writing code

# NYTimes Top Stories
```{r}

baseurl <- "https://api.nytimes.com/svc/topstories/v2/home.json?"
limit <- 20
url <- paste0(baseurl, "&api-key=", apikey) 

# Issue the query:
nyt.json <- fromJSON(file = url)

class(nyt.json)


nyt.json$status
summary(nyt.json)
length(nyt.json$results)
names(nyt.json$results[[1]])
```

Now Instantiating a nyt dataframe to store json response 
```{r}
nyt.data.df <- data.frame(
    id=numeric(),
    section = character(),
    title=character(),
    url= character(),
    author = character(),
    published_date =character()
  )

for(i in 1:length(nyt.json$results)){    
      id <- as.numeric(i)
      section <- nyt.json$results[[i]]$section
      title <- nyt.json$results[[i]]$title
      url <- nyt.json$results[[i]]$url
      author<- nyt.json$results[[i]]$byline
      published_date <- nyt.json$results[[i]]$published_date
           nyt.data.df <- rbind(nyt.data.df,(cbind(id,section,title,url,author,published_date)),stringsAsFactors = F)
}


DT::datatable(nyt.data.df, options = list(pageLength = 5))

```

# EDA -  NYTimes Top Stories

```{r echo=FALSE,warning=FALSE, cache=FALSE, message=FALSE}
library(ggplot2)
ggplot(nyt.data.df,aes(x=section))+
  geom_bar(aes(section, fill = ..count..))+
  labs(title = "NYTimes Article: Top Stories", 
                         x = "Article Section", 
                         y = "Total Count") +
  theme_bw()+
theme(axis.text.x = element_text(angle = 65, hjust = 1))

```



