---
title: "DATA607 Week 5 Assignment"
author: "Keshaw Sahay"
date: "February 29, 2020"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    number_section:  true
    theme: cerulean
    highlight:  tango
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. The chart above describes arrival delays for two airlines across five destinations. Your task is to: <br>
**Task#1: **  Create a .CSV file (or optionally, a MySQL database!) that includes all of the information above. You’re encouraged to use a “wide” structure similar to how the information appears above, so that you can practice tidying and transformations as described below. <br>
**Task#2: ** Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy and transform your data. <br>
**Task#3: ** Perform analysis to compare the arrival delays for the two airlines. <br>
**Task#4: ** Your code should be in an R Markdown file, posted to rpubs.com, and should include narrative <br> descriptions of your data cleanup work, analysis, and conclusions. Please include in your homework submission:

# Data Acquisition

```{r}
library(knitr)
flights.data = read.csv('https://raw.githubusercontent.com/keshaws/CUNY_MSDS_2020/master/DATA607/fligthts.csv',sep = ',')
kable(flights.data)
```

# Environment Set Up

```{r warning=FALSE, message=FALSE}
library(tidyr)
library(dplyr)
```

```{r}
summary(flights.data)
```

# Tidying Data

The dataset is untidy, so using gather function to convert the imported data into a 'tidy' format

```{r}
flight <- gather(flights.data, City, Flight_Count, Los.Angeles:Seattle)
kable(flight)
```

#_Total number of flights by each carrier (airlines) to different destinations (Cities)_

```{r}
total.flights <- flight %>% group_by(Carrier) %>% 
      summarize(Total_Flights = sum(Flight_Count))
total.flights
```

#_On-time flights for each carrier_

```{r}
flights.ontime <- flight %>% group_by(Carrier) %>% 
  filter(Status == 'on_time') %>% 
  summarize(FlightsOnTime = sum(Flight_Count))
flights.ontime
```

#_Delayed flights for each carrier_

```{r}
flights.delayed <- flight %>% group_by(Carrier) %>% 
  filter(Status == 'delayed') %>% 
  summarize(FlightsDelayed = sum(Flight_Count))
flights.delayed
```

```{r}
flights.info = cbind(flights.ontime,FlightsDelayed = flights.delayed$FlightsDelayed, TotalFlights=total.flights$Total_Flights)
flights.info
```

```{r}
flights.summary <- mutate(flights.info, PercentOnTime = FlightsOnTime/TotalFlights, PercentDelayed= FlightsDelayed/TotalFlights)
flights.summary
```

# Analysis

## EDA (Exploratory Data Analysis)

```{r}
Carrier.Alaska <- filter(flight, Carrier == 'Alaska')
kable(Carrier.Alaska)
```

```{r}
Carrier.AMWEST <- filter(flight, Carrier == 'AMWEST')
kable(Carrier.AMWEST)
```

For Alaskan airlines which city had most number of delays:

```{r}
alaska.delays <- filter(Carrier.Alaska,Status=='delayed')
city.alaska.delays <- alaska.delays %>%
                      arrange(desc(alaska.delays$Flight_Count))
kable(city.alaska.delays)
```

```{r}
alaska.ontime <- filter(Carrier.Alaska,Status=='on_time')
city.alaska.ontime.rank <- alaska.ontime %>%
                      arrange(desc(alaska.ontime$Flight_Count))
kable(city.alaska.ontime.rank)
```

```{r}
city.amwest.delays <-  filter(Carrier.AMWEST, Status == 'delayed')
city.amwest.delays <- city.amwest.delays %>%
                      arrange(desc(city.amwest.delays$Flight_Count))
kable(city.amwest.delays)
```



```{r}
city.flight.delays <-  filter(flight, Status == 'delayed')
city.flight.delays <- city.flight.delays %>%
                    arrange(desc(city.flight.delays$Flight_Count))

```


```{r}
kable(city.flight.delays)
```


```{r fig1, fig.height = 6, fig.width = 8}
library(ggplot2)
ggplot(data = city.flight.delays, aes(x=Carrier,y=Flight_Count))+
  #geom_point(alpha = 0.5, size = 5, color ='blue')
  geom_bar(stat = 'identity',aes(fill=Carrier))+
  geom_text(aes(x = Carrier, y = Flight_Count, 
                                    label = paste(Flight_Count),
                                    group = Carrier,
                                    vjust = -0.4)) +
  labs(title = "Comparing Delays by Carrier & City", 
                         x = "Carrier", 
                         y = "Delay Flight Count") +
  facet_wrap(~City, ncol = 5)+
  theme_bw()
```

spread function to view the dataset for delayed flights

```{r}
flight.delay.spread <- spread(city.flight.delays,City, Flight_Count)
```

```{r}
kable(flight.delay.spread)
```


```{r}
flight.df <- flight %>%
    spread(Status,'Flight_Count')
kable(flight.df)
```


```{r}
flight_percentile <- mutate(flight.df,delay_percent=round((delayed/(delayed+on_time)*100),2),
                            onTime_percent=round((on_time/(delayed+on_time))*100,2))
kable(flight_percentile)
```



```{r fig2, fig.height = 6, fig.width = 8}
library(ggplot2)
ggplot(data = flight_percentile, aes(x=Carrier,y=delay_percent))+
  geom_bar(stat = 'identity',aes(fill=Carrier))+
  geom_text(aes(x = Carrier, y = delay_percent, 
                                    label = paste(delay_percent,'%'),
                                    group = Carrier,
                                    vjust = -0.4)) +
  labs(title = "Comparing Delays(%) by Carrier & City", 
                         x = "Carrier", 
                         y = "Delay Percent") +
  facet_wrap(~City, ncol = 5)+
  theme_bw()
```




```{r fig3, fig.height = 6, fig.width = 8}
library(ggplot2)
ggplot(data = flight_percentile, aes(x=Carrier,y=onTime_percent))+
  geom_bar(stat = 'identity',aes(fill=Carrier))+
  geom_text(aes(x = Carrier, y = onTime_percent, 
                                    label = paste(onTime_percent,'%'),
                                    group = Carrier,
                                    vjust = -0.4)) +
  labs(title = "Comparing OnTime(%) by Carrier & City", 
                         x = "Carrier", 
                         y = "OnTime Percent") +
  facet_wrap(~City, ncol = 5)+
  theme_bw()
```


# Conclusion:
 + San Fransicso has highest delay rate for both carriers
 + Phoenix has highest on time rate for both carriers
 + AMWEST has higher delay rate than Alaska across 5 cities

